
Seeding a GIS with R notebook.

First touched on this from the R 251 class.
see R_phw251_bonus_problem_set.html

#ref choropleth

**^ tin zink ~/tin-gh/PHW251_Fall2022/problem sets/problem set bonus ^**>  wc problem_set_bonus.Rmd 

```{r, setup }

library(pacman)
p_load(tidyverse)
p_load(plotly)

```


# Reference materials

* (sf cheat sheet)[https://github.com/r-spatial/sf#cheatsheet] 


```{r, plotly_web_cholopleth_eg, include=T }

# ref example from plotly web site 
# https://plotly.com/r/choropleth-maps/

df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
df$hover <- with(df, paste(state, '<br>', "Beef", beef, "Dairy", dairy, "<br>",
                           "Fruits", total.fruits, "Veggies", total.veggies,
                           "<br>", "Wheat", wheat, "Corn", corn))

fig <- plot_geo(df, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~total.exports, text = ~hover, locations = ~code,
    color = ~total.exports, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "Millions USD")
fig <- fig %>% layout(
    title = '2011 US Agriculture Exports by State<br>(Hover for breakdown)'
  )

fig
```

PHW251_bonus_problem_set_Q14
(https://github.com/tin6150/PHW251_Fall2022/blob/main/problem%20sets/problem%20set%20bonus/problem_set_bonus.Rmd)[https://github.com/tin6150/PHW251_Fall2022/blob/main/problem%20sets/problem%20set%20bonus/problem_set_bonus.Rmd]


```{r gis_choropleth, include=T}

suicide_CSV_file="https://raw.githubusercontent.com/tin6150/PHW251_Fall2022/main/problem%20sets/problem%20set%20bonus/data/Suicide%20Mortality%20by%20State.csv"

# data pulled from CDC website described above 
#df_suicide <- read_csv("data/Suicide Mortality by State.csv") %>%
df_suicide <- read_csv(suicide_CSV_file) %>%
    filter(YEAR == 2018)


## add hover pop up text as extra col of data to DF
df_suicide$hover = with(
  df_suicide,
  paste( NAME, '<br>',
         'Death Rate: ', RATE,    '<br>',
         'Death: '     , DEATHS,  '<br>'
         ) 
)

map_q14 = plot_ly(df_suicide,
        type="choropleth",
        locationmode = "USA-states") %>%
  layout(geo=list(scope="usa"))   %>% add_trace(
    z            = ~RATE
    , text       = ~hover
    , locations  = ~STATE
    , color      = ~RATE
    , colors     = 'Purples'
  )  %>%   
  layout(      
    title = paste(
      "Suicide Mortality by State in 2018"
      , "<br> The number of deaths per 100,000 total population"
) )


map_q14 # map renders well in Rstudio, but not to pdf.


```

xref:  ~/Dropbox/mph_hw/gis_phw272a/example_index.Sn50.txt  for many gis related constructs from intro GIS in R class

PHW 292A - GIS intro with R
Week 1: Introduction to Geographic Information Systems (GIS) for Public Health
January 10–15

https://bcourses.berkeley.edu/courses/1521079/pages/week-1

```{r gis272A_wk1}

p_load( dplyr )
p_load( ggplot2 )
# sf need below, installed on Zink
# sudo apt install libudunits2-dev
p_load( sf )
p_load( sp )
p_load( tigris )
p_load( geodata )
p_load( raster )
p_load( paletteer ) # this is the fixed, corrected package name

# from lab1, clever pacman

package_list = c("ggplot2", "dplyr", "sf", "rgdal", "raster", "here", "scales")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(package_list, character.only=TRUE)


# load shapefile, display it 
wk1_ca_shape_file_path = "/home/tin/Dropbox/mph_hw/gis_phw272a/wk1/lab1/CA_Counties/CA_Counties_TIGER2016.shp"

## x11(type = "cairo")        ## use this to create display if not using Rmd notebook

wk1_ca_shpSn50 = st_read( ca_shape_file_path )
class( wk1_ca_shpSn50 )
ggplot() + geom_sf( data=wk1_ca_shpSn50 )

```


```{r gis272A_wk1_interpretation_assignment}

## Part 5 page 11 of wk1 interpretation assignment
## file:///home/tin/Dropbox/mph_hw/gis_phw272a/wk1/wk1_gis272a.InterpretationAssignment_SpatialDataFrames.pdf

p_load( tigris )

# tigris: https://www.rdocumentation.org/packages/tigris/versions/1.6.1
# download TIGER/Line shape file directly (US Census)
# internet connection req, data is not "installed" as part of tigris pkg install
# but cached for future use

# tracts are 1:500k 
wk1_assign_shp_Bay = 
  tracts( state='CA',
          county = c("Alameda", "Contra Costa", "Marin",
                     "Napa", "San Francisco", "San Mateo",
                     "Santa Clara", "Solano", "Sonoma")  
          )

# counties are 1:500k; 1:5m; 1:20m
wk1_assign_shp_CA = counties( state="CA" )

# many other features can be downloaded, regions, places, zctas, area_water, urban_areas, native_areas, rails...

class(wk1_assign_shp_Bay)  # sf
class(wk1_assign_shp_CA)   # sf

# several way to plot shape data
ggplot( wk1_assign_shp_Bay ) + geom_sf() 

ggplot() + geom_sf( data=wk1_assign_shp_CA )

ggplot( wk1_assign_shp_Bay ) + geom_sf() + geom_sf( data=wk1_assign_shp_CA )
ggplot() + geom_sf( data=wk1_assign_shp_Bay ) + geom_sf( data=wk1_assign_shp_CA )
# either of the above lines, the 2nd geom_sf() data overwrite the 1st load.  

# ah, but plot state first, then bay area tracts, work!
ggplot() +  geom_sf( data=wk1_assign_shp_CA ) + geom_sf( data=wk1_assign_shp_Bay )

```
